---
title: "ABG200A Statistics and Data Science"
author: "Hao Cheng"
date: "10/7/2020"
output: html_document
---

```{r setup1, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Exercise 1: Creative Writing Study

* Experimental Units: writers
* Treatment: type of questionnaire given at the beginning of the study (questions emphasize either intrinsic or extrinsic rewards)
* Random assignment: 24 intrinsic, 23 extrinsic
* Response: average of 12 evaluations of creativity on 40 points scale

```{r}
intrinsic = c(12.0, 12.0, 12.9, 13.6, 16.6, 17.2,
              17.5, 18.2, 19.1, 19.3, 19.8, 20.3,
              20.5, 20.6, 21.3, 21.6, 22.1, 22.2,
              22.6, 23.1, 24.0, 24.3, 26.7, 29.7)
extrinsic = c( 5.0,  5.4,  6.1, 10.9, 11.8, 12.0,
              12.3, 14.8, 15.0, 16.8, 17.2, 17.2,
              17.4, 17.5, 18.5, 18.7, 18.7, 19.2,
              19.5, 20.7, 21.2, 22.1, 24.0)

nintrinsic = length(intrinsic)
boxplot(intrinsic,extrinsic)
```

### Model-based Inference
```{r}
t.test(intrinsic, extrinsic, var.equal = TRUE)
```

#### model diagnostics

#### homogeneous variances
```{r}
var.test(intrinsic,extrinsic)
```

#### normality (1)
```{r}
plot(density(intrinsic))
plot(density(extrinsic))

library(car)
qqPlot(intrinsic)
```

#### normality (2)
```{r}
nquantiles = nintrinsic
quantile1  = 1/nquantiles
qs         = numeric(nquantiles)
for (i in 1:nquantiles){
  qi     = quantile1*i
  qs[i]  = qnorm(qi)
}
plot(qs,intrinsic)
```


### Randomization (design-based inference)
```{r}
nsamples   = 10000
diffs      = numeric(nsamples)
for (i in 1:nsamples){
  randi      = sample(c(extrinsic,intrinsic))
  intrinsici = randi[1:nintrinsic]
  extrinsici = randi[-(1:nintrinsic)]
  diffs[i]   = mean(intrinsici) - mean(extrinsici)
}
diff_obs = mean(intrinsic) - mean(extrinsic)

hist(diffs,col="gray",las=1,main="")
abline(v=diff_obs,col="red")

(mean(diffs >= diff_obs))
```

## Exercise 2: Movie Ratings

|            |movie 1|movie 2|movie 3|
|:----------:|:-----:|:-----:|:-----:|
| customer 1 |  4    |   1   |       |
| customer 1 |       |   3   |   5   |
| customer 1 |       |       |   3   |
| customer 1 |  3    |   1   |       |

### Data
```{r}
customer= factor(c(1,1,2,2,3,4,4))
movie   = factor(c(1,2,2,3,3,1,2))
y       = c(4,1,3,5,3,3,1)
d       = data.frame(customer,movie,y)
d
```

### estimable function (1)

```{r}
library(MASS)

y = c(4,1,3,5,3,3,1)
X = matrix(c(
1,1,0,0,0,1,0,0,
1,1,0,0,0,0,1,0,
1,0,1,0,0,0,1,0,
1,0,1,0,0,0,0,1,
1,0,0,1,0,0,0,1,
1,0,0,0,1,1,0,0,
1,0,0,0,1,0,1,0
),byrow=T,nrow=7)

bhat=ginv(t(X)%*%X)%*%t(X)%*%y

bhat
```

### estimable function (2)

```{r}
o=lm(y~customer+movie,data=d)
summary(o)
model.matrix(o)
coef(o)
fitted(o)
resid(o)
```

### estimable function (3)

```{r}
o=lm(y~customer+movie+customer*movie,data=d)
summary(o)
model.matrix(o)
coef(o)
fitted(o)
```




